Hydrodynamics and rheology of a vesicle doublet suspension

### 2019-07-15 vesicle_code/docs/adhesion/adhesion.tex
added a small todo
working on revisions

Collective hydrodynamics of vesicles is dictated by the vesicle-vesicle
interactions.  In a quiescent flow, vesicle-vesicle adhesion leads to
the formation of vesicle doublets~\cite{Ziherl2007_PRL,
ZiherlSvetina2007_PNAS} or
clusters~\cite{SvetinaZiherl2008_Bioelectrochemistry,GuWangGunzburger2016_MathBiol,
FlormannAouane2017_SciReports}.  As a model for red blood cell (RBC)
aggregates, a simplified model for adhesive vesicle-vesicle interactions
can reproduce vesicle shapes similar to those observed in
experiments of fibrinogen-induced RBC
aggregates~\cite{SvetinaZiherl2008_Bioelectrochemistry,GuWangGunzburger2016_MathBiol,
FlormannAouane2017_SciReports,HooreYayaPodgorski2018_SoftMatt}.  Using the Lennard-Jones
(L.-J.)~potential for point-point interaction between two vesicle
membranes without the molecular details of adhesive interactions between
RBCs, Flormann {\em et al.}~\cite{FlormannAouane2017_SciReports} found
that, under strong adhesion, 
the membrane may buckle to a sigmoidal shape in the contact region.
Such a sigmoidal vesicle shape is 
also observed in RBC doublets and explained in a slightly different
model~\cite{ZiherlSvetina2007_PNAS}.  It remains unknown how adhesive
interactions between vesicles might lead to different vesicle
hydrodynamics that results in different rheological properties of a
vesicle suspension. Such studies will provide useful insight into the
rheology of blood flow and how to use nano solutes to control the
hypertension by adjusting the adhesive interactions between RBCs.

@@

In a quiescent environment, sub-micron size vesicles are found to form a
doublet due to their van der Waals' attractive
interactions~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}.
Due to the strong van der Waals' adhesive force, the vesicles are far
from spherical shape and the membrane is almost flat in the ``contact"
region. Gires {\em et al.}~used small-deformation analysis to
investigate the hydrodynamic interactions between two vesicles in a
planar shear flow with a long separation
distance~\cite{GiresDankerMisbah2012_PRE}. They found that the vesicle
interaction could be either attractive or repulsive depending on the
organization of the two vesicles relative to the shear flow. To the best
of authors' knowledge, the effects of close-range vesicle adhesion on
their hydrodynamics under an external flow have not been studied and
quantified. The goal of this work is to use state-of-the-art boundary
integral simulations to numerically investigate the dynamics of two
vesicles in both planar shear flow and extensional flow for a wide range
of vesicle shapes and adhesive strength and distance.

@@

\begin{table}[htp]
\begin{minipage}{0.45\textwidth}
  $\mathcal{H} = 1.5$ \\
  \begin{tabular}{>{\centering\arraybackslash}p{2cm}
>{\centering\arraybackslash}p{2cm}
>{\centering\arraybackslash}p{2cm}}
  $\chi$ & $p$ & $\lambda_{\mathrm{eff}}$ \\
  \hline
%  0.50 & 38.06 & 2.92 \\
%  0.75 & 24.81 & 2.85 \\
%  1.00 & 19.47 & 3.00 
  0.50 & 36.34 & 2.49 \\
  0.75 & 25.86 & 2.72 \\
  1.00 & 20.73 & 2.96 
\end{tabular}
\end{minipage}
%\hfill
\begin{minipage}{0.45\textwidth}
  $\mathcal{H} = 3.5$ \\
  \begin{tabular}{>{\centering\arraybackslash}p{2cm}
>{\centering\arraybackslash}p{2cm}
>{\centering\arraybackslash}p{2cm}}
  $\chi$ & $p$ & $\lambda_{\mathrm{eff}}$ \\
  \hline
  0.50 & 28.46 & 1.66 \\
  0.75 & 22.01 & 2.17 \\
  1.00 & 17.91 & 2.44 
\end{tabular}
\end{minipage}
\caption{\label{tbl:aspectRatio} The shear rate, period, and effective
aspect ratio of a vesicle doublet with reduced area $\Delta A = 0.9$,
separation distance $\delta = 0.4$, and 
Hamaker constants $\mathcal{H} = 1.5$ and $\mathcal{H} = 3.5$.  The
effective aspect ratios depend on both the Hamaker constant and the
shear rate, indicating that the reported Hamaker constants, the dynamics
of the doublet differ from that of a Jeffery orbit.}
\end{table}
\todo[inline]{period results with $\delta$ = 0.1, 0.2, 0.3, 0.4 so that
we can simply say how it affects the period}

@@

Finally, we investigate the rheology of a suspension of a doublet by
computing the effective viscosity of a doublet and compare it to the
effective viscosity of a single tank-treading vesicle. The effective
viscosity is defined as the viscosity of a homogeneous Newtonian fluid
with the same energy dissipation per macroscopic element of fluid.  In a
simple shear flow, the intrinsic viscosity, $[\mu]$ is
\begin{align*}
  [\mu]:= \frac{\mu_{\mathrm{eff}} - \mu_0}{\phi \mu_0} = 
  \frac{1}{\chi \mu_0 (T_e - T_i)} \int_{T_i}^{T_e} 
  \langle \tau_{12} \rangle dt,
\end{align*}
where
\begin{align*}
  \langle \tau \rangle = \frac{1}{|\omega|} \int_{\gamma}
    \xxi \otimes \xx ds,
\end{align*}
$\phi$ is the area fraction of vesicles, $\tau$ is the stress due to the
vesicles, $\langle \cdot \rangle$ is the spatial average, $|\omega|$ is
the total area of both vesicles, and $\xxi$ is the traction as defined
in equation~\eqref{eqn:traction}.  In
Figure~\ref{fig:shearIntrinsicViscosity}(a), we compare the intrinsic
viscosity of a single tank-treading vesicle to a doublet with Hamaker
constant $\mathcal{H} = 0.7$, separation distance $\delta = 0.4$, and
shear rate $\chi = 0.5$.  To validate our simulations, we superimpose
(black marks) the intrinsic viscosity calculated by Ghigliotti {\em et
al.}~\cite{GhigliottiBibenMisbah2010_JFM}.  The presence of the doublet
slightly increases the intrinsic viscosity at all the reduced areas.
We also compute the intrinsic viscosity of a doublet with $\Delta A =
0.9$, $\chi = 0.5$, and $\mathcal{H} = 2.1$.  Again, this doublet has a
similar period to the case $\mathcal{H} = 2.1$.  At this larger Hamaker
constant, the effective viscosity increases slightly from $1.97$ to
$2.06$.

@@

BQ and YNY contributed equally to model development, code development,
and simulations.  All authors contributed equally to writing the
manuscript.

@@

\bibliography{refs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{\label{sec:level2}Second-level heading: Formatting}
%
%This file may be formatted in either the \texttt{preprint} or
%\texttt{reprint} style. \texttt{reprint} format mimics final journal output. 
%Either format may be used for submission purposes. \texttt{letter} sized paper should
%be used when submitting to APS journals.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsubsection{Wide text (A level-3 head)}
%The \texttt{widetext} environment will make the text the width of the
%full page, as on page~\pageref{eq:wideeq}. (Note the use the
%\verb+\pageref{#1}+ command to refer to the page number.) 
%\paragraph{Note (Fourth-level head is run in)}
%The width-changing commands only take effect in two-column formatting. 
%There is no effect if text is in a single column.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{\label{sec:citeref}Citations and References}
%
%
%
%A citation in text uses the command \verb+\cite{#1}+ or
%\verb+\onlinecite{#1}+ and refers to an entry in the bibliography. 
%An entry in the bibliography is a reference to another document.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsubsection{Citations}

%Because REV\TeX\ uses the \verb+natbib+ package of Patrick Daly, 
%the entire repertoire of commands in that package are available for your document;
%see the \verb+natbib+ documentation for further details. Please note that
%REV\TeX\ requires version 8.31a or later of \verb+natbib+.
%
%\paragraph{Syntax}
%The argument of \verb+\cite+ may be a single \emph{key}, 
%or may consist of a comma-separated list of keys.
%The citation \emph{key} may contain 
%letters, numbers, the dash (-) character, or the period (.) character. 
%New with natbib 8.3 is an extension to the syntax that allows for 
%a star (*) form and two optional arguments on the citation key itself.
%The syntax of the \verb+\cite+ command is thus (informally stated)
%\begin{quotation}\flushleft\leftskip1em
%\verb+\cite+ \verb+{+ \emph{key} \verb+}+, or\\
%\verb+\cite+ \verb+{+ \emph{optarg+key} \verb+}+, or\\
%\verb+\cite+ \verb+{+ \emph{optarg+key} \verb+,+ \emph{optarg+key}\ldots \verb+}+,
%\end{quotation}\noindent
%where \emph{optarg+key} signifies 
%\begin{quotation}\flushleft\leftskip1em
%\emph{key}, or\\
%\texttt{*}\emph{key}, or\\
%\texttt{[}\emph{pre}\texttt{]}\emph{key}, or\\
%\texttt{[}\emph{pre}\texttt{]}\texttt{[}\emph{post}\texttt{]}\emph{key}, or even\\
%\texttt{*}\texttt{[}\emph{pre}\texttt{]}\texttt{[}\emph{post}\texttt{]}\emph{key}.
%\end{quotation}\noindent
%where \emph{pre} and \emph{post} is whatever text you wish to place 
%at the beginning and end, respectively, of the bibliographic reference
%(see Ref.~[\onlinecite{witten2001}] and the two under Ref.~[\onlinecite{feyn54}]).
%(Keep in mind that no automatic space or punctuation is applied.)
%It is highly recommended that you put the entire \emph{pre} or \emph{post} portion 
%within its own set of braces, for example: 
%\verb+\cite+ \verb+{+ \texttt{[} \verb+{+\emph{text}\verb+}+\texttt{]}\emph{key}\verb+}+.
%The extra set of braces will keep \LaTeX\ out of trouble if your \emph{text} contains the comma (,) character.
%
%The star (*) modifier to the \emph{key} signifies that the reference is to be 
%merged with the previous reference into a single bibliographic entry, 
%a common idiom in APS and AIP articles (see below, Ref.~[\onlinecite{epr}]). 
%When references are merged in this way, they are separated by a semicolon instead of 
%the period (full stop) that would otherwise appear.
%
%\paragraph{Eliding repeated information}
%When a reference is merged, some of its fields may be elided: for example, 
%when the author matches that of the previous reference, it is omitted. 
%If both author and journal match, both are omitted.
%If the journal matches, but the author does not, the journal is replaced by \emph{ibid.},
%as exemplified by Ref.~[\onlinecite{epr}]. 
%These rules embody common editorial practice in APS and AIP journals and will only
%be in effect if the markup features of the APS and AIP Bib\TeX\ styles is employed.
%
%\paragraph{The options of the cite command itself}
%Please note that optional arguments to the \emph{key} change the reference in the bibliography, 
%not the citation in the body of the document. 
%For the latter, use the optional arguments of the \verb+\cite+ command itself:
%\verb+\cite+ \texttt{*}\allowbreak
%\texttt{[}\emph{pre-cite}\texttt{]}\allowbreak
%\texttt{[}\emph{post-cite}\texttt{]}\allowbreak
%\verb+{+\emph{key-list}\verb+}+.

>>>>

Collective hydrodynamics of vesicles is dictated by the vesicle-vesicle
interactions.  In a quiescent flow, vesicle-vesicle adhesion leads to
the formation of vesicle doublets~\cite{Ziherl2007_PRL,
ZiherlSvetina2007_PNAS} or
clusters~\cite{SvetinaZiherl2008_Bioelectrochemistry,GuWangGunzburger2016_MathBiol,
FlormannAouane2017_SciReports}.  As a model for red blood cell (RBC)
aggregates, a simplified model for adhesive vesicle-vesicle interactions
can reproduce vesicle shapes similar to those observed in
experiments of fibrinogen-induced RBC
aggregates~\cite{SvetinaZiherl2008_Bioelectrochemistry,GuWangGunzburger2016_MathBiol,
FlormannAouane2017_SciReports,HooreYayaPodgorski2018_SoftMatt}.  Using the Lennard-Jones
(L.-J.)~potential for point-point interaction between two vesicle
membranes without the molecular details of adhesive interactions between
RBCs, Flormann {\em et al.}~\cite{FlormannAouane2017_SciReports} found
that, under strong adhesion, 
the membrane may buckle to a sigmoidal shape in the contact region.
Such a sigmoidal vesicle shape is 
also observed in RBC doublets and explained in a slightly different
model~\cite{ZiherlSvetina2007_PNAS}.  It remains unknown how adhesive
interactions between vesicles might lead to different vesicle
hydrodynamics that results in different rheological properties of a
vesicle suspension. Such studies will provide useful insight into the
rheology of blood flow~\cite{chi-usa-del-gre-nan-gue1967} and how to use
nano solutes to control the hypertension by adjusting the adhesive
interactions between RBCs.
\todo[inline]{Add references by Brust aet al.~ and Claveria et al}

@@

In a quiescent environment, sub-micron size vesicles are found to form a
doublet due to their van der Waals' attractive
interactions~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}.
Due to the strong van der Waals' adhesive force, the vesicles are far
from spherical shape and the membrane is almost flat in the ``contact"
region. Gires {\em et al.}~used small-deformation analysis to
investigate the hydrodynamic interactions between two vesicles in a
planar shear flow with a long separation
distance~\cite{GiresDankerMisbah2012_PRE}.  \todo[inline]{Also cite
Gires et al.~Physics of Fluids 26, 013304, 2014} They found that the
vesicle interaction could be either attractive or repulsive depending on
the organization of the two vesicles relative to the shear flow. To the
best of authors' knowledge, the effects of close-range vesicle adhesion
on their hydrodynamics under an external flow have not been studied and
quantified. The goal of this work is to use state-of-the-art boundary
integral simulations to numerically investigate the dynamics of two
vesicles in both planar shear flow and extensional flow for a wide range
of vesicle shapes and adhesive strength and distance.

@@

\begin{table}[htp]
\begin{minipage}{0.45\textwidth}
  $\mathcal{H} = 1.5$ \\
  \begin{tabular}{>{\centering\arraybackslash}p{2cm}
>{\centering\arraybackslash}p{2cm}
>{\centering\arraybackslash}p{2cm}}
  $\chi$ & $p$ & $\lambda_{\mathrm{eff}}$ \\
  \hline
%  0.50 & 38.06 & 2.92 \\
%  0.75 & 24.81 & 2.85 \\
%  1.00 & 19.47 & 3.00 
  0.50 & 36.34 & 2.49 \\
  0.75 & 25.86 & 2.72 \\
  1.00 & 20.73 & 2.96 
\end{tabular}
\end{minipage}
%\hfill
\begin{minipage}{0.45\textwidth}
  $\mathcal{H} = 3.5$ \\
  \begin{tabular}{>{\centering\arraybackslash}p{2cm}
>{\centering\arraybackslash}p{2cm}
>{\centering\arraybackslash}p{2cm}}
  $\chi$ & $p$ & $\lambda_{\mathrm{eff}}$ \\
  \hline
  0.50 & 28.46 & 1.66 \\
  0.75 & 22.01 & 2.17 \\
  1.00 & 17.91 & 2.44 
\end{tabular}
\end{minipage}
\caption{\label{tbl:aspectRatio} The shear rate, period, and effective
aspect ratio of a vesicle doublet with reduced area $\Delta A = 0.9$,
separation distance $\delta = 0.4$, and 
Hamaker constants $\mathcal{H} = 1.5$ and $\mathcal{H} = 3.5$.  The
effective aspect ratios depend on both the Hamaker constant and the
shear rate, indicating that the reported Hamaker constants, the dynamics
of the doublet differ from that of a Jeffery orbit.}
\end{table}

@@

Finally, we investigate the rheology of a suspension of a doublet by
computing the effective viscosity of a doublet and compare it to the
effective viscosity of a single tank-treading vesicle. The effective
viscosity is defined as the viscosity of a homogeneous Newtonian fluid
with the same energy dissipation per macroscopic element of fluid.  In a
simple shear flow, the intrinsic viscosity, $[\mu]$ is
\begin{align*}
  [\mu]:= \frac{\mu_{\mathrm{eff}} - \mu_0}{\phi \mu_0} = 
  \frac{1}{\chi \mu_0 (T_e - T_i)} \int_{T_i}^{T_e} 
  \langle \tau_{12} \rangle dt,
\end{align*}
where
\begin{align*}
  \langle \tau \rangle = \frac{1}{|\omega|} \int_{\gamma}
    \xxi \otimes \xx ds,
\end{align*}
$\phi$ is the area fraction of vesicles, $\tau$ is the stress due to the
vesicles, $\langle \cdot \rangle$ is the spatial average, $|\omega|$ is
the total area of both vesicles, and $\xxi$ is the traction as defined
in equation~\eqref{eqn:traction}.  In
Figure~\ref{fig:shearIntrinsicViscosity}(a), we compare the intrinsic
viscosity of a single tank-treading vesicle to a doublet with Hamaker
constant $\mathcal{H} = 0.7$, separation distance $\delta = 0.4$, and
shear rate $\chi = 0.5$.  To validate our simulations, we superimpose
(black marks) the intrinsic viscosity calculated by Ghigliotti {\em et
al.}~\cite{GhigliottiBibenMisbah2010_JFM}.  The presence of the doublet
slightly increases the intrinsic viscosity at all the reduced areas.
We also compute the intrinsic viscosity of a doublet with $\Delta A =
0.9$, $\chi = 0.5$, and $\mathcal{H} = 2.1$.  Again, this doublet has a
similar period to the case $\mathcal{H} = 0.7$. At this larger Hamaker
constant, the effective viscosity increases slightly from $1.97$ to
$2.06$.

@@

BQ and YNY contributed equally to model development and simulations. BQ
developed the code. All authors contributed equally to writing the
manuscript.

@@

\bibliography{refs}
